---
title: "BEM_RightingResponse"
author: "Kevin Mack"
date: "2/18/2016"
output: html_document
---

Preliminary Analysis
--
This is preliminary/exploratory analysis of sublethal righting responses of _P. armatus_. I ran an ANCOVA and on the main effects: site, sex, and block with crab size (CW) as a covariate. A significant block effect was detected. This prelimary analysis indicates that the covariate CW was the strongest effect, however CW is closely linked to crab sex, with males generally being larger than females. There were no significant interactions. This analysis informed the analysis of main effects conducted next.
```{r}
KM_Righting <- read.csv("~/Applied Quant Project/Righting-Response/Trial1_sublethal.csv")
KM_Righting$Block=as.factor(KM_Righting$Block)
library(survival)

#ANCOVA of main effests: Righting Response
options(contrasts=c("contr.sum", "contr.poly"))
fit1=(aov(lm(Duration~Site*Sex*Block+CW, data=KM_Righting)))
drop1(fit1,~.,test='F')

#Boxplots: Righting Response
plot(KM_Righting$Sex, KM_Righting$Duration)
plot(KM_Righting$Site, KM_Righting$Duration)
plot(KM_Righting$CW,KM_Righting$Duration)
plot(KM_Righting$Duration, KM_Righting$CW ) 
#closley related to sex. Examine males and females separetely?

```


Analysis of Main Effects
--
This is analysis of righting responses of _P. armatus_. I tested the main effects: site, sex, block and crab size (CW) by generating Kaplan-Meier curves for righting response data, main effects were tested with a Log Rank test. I next fitted these curves to a Cox proportional hazards model to determine differences in hazard risk between main effects. No significant differences were found between sites, but were detected between sexes and sizes. This helped me determine how to present differences between my main effects.
```{r}
#KM Curves of main effects
rr_mod1=survfit(Surv(Duration,Status)~Block+Site+Sex, data=KM_Righting)
plot(rr_mod1, ylab='Survival Probability', xlab='Duration') 
 
#Log Rank test of main effects tests for differences
rr_mod1_dif=survdiff(Surv(Duration,Status)~Block+Site+Sex, data=KM_Righting)
rr_mod1_dif

#Cox proportional hazards model fit
rr_mod1_cox=coxph(Surv(Duration,Status)~Block+Site+Sex+CW, data=KM_Righting)
summary(rr_mod1_cox)

#ANOVA of main effects, Useful for checking results 
anova(rr_mod1_cox)
```


Analysis of Spatial Scale
--
The following is analysis of righting responses of _P. armatus_ at three spatial scales. I tested the main effects, grouping data by site, location, and region, to determine inter-site and region differences in righting response. Differences in righting response exist between sites, locations, and regions, but can be best summarized on a regional scale (North vs South)
```{r}
##Analysis by SITE
#KM Curves of main effects
site_righting=survfit(Surv(Duration,Status)~Block+Site+Sex, data=KM_Righting)
plot(site_righting, ylab='survival probabilty', xlab='Duration', main='Site') 

#Log Rank test of main effects tests for differences
site_righting_dif=survdiff(Surv(Duration,Status)~Block+Site+Sex,
                           data=KM_Righting)
site_righting_dif

#Cox proportional hazards model fit
site_righting_cox=coxph(Surv(Duration,Status)~Block+Site+Sex+CW, 
                        data=KM_Righting)
summary(site_righting_cox)

#ANCOVA of main effects 
anova(site_righting_cox)
```

```{r}
##Analysis by LOCATION
#KM Curves of main effects
location_righting=survfit(Surv(Duration,Status)~Block+Location+Sex, 
                          data=KM_Righting)
plot(location_righting, ylab='righting probabilty', xlab='Duration', 
     main='Location')

#Log Rank test of main effects tests for differences
location_righting_dif=survdiff(Surv(Duration,Status)~Block+Location+Sex,
                               data=KM_Righting)
location_righting_dif

#Cox proportional hazards model fit
location_righting_cox=coxph(Surv(Duration,Status)~Block+Location+Sex+CW,
                            data=KM_Righting)
summary(location_righting_cox)

#ANCOVA of main effects 
anova(location_righting_cox)
```

```{r}
##Analysis by REGION
#KM Curves of main effects
region_righting=survfit(Surv(Duration,Status)~Block+Region+Sex, 
                        data=KM_Righting)
plot(region_righting, ylab='survival probabilty', xlab='Duration', 
     main='Region')

#Log Rank test of main effects tests for differences
region_righting_dif=survdiff(Surv(Duration,Status)~Block+Region+Sex, 
                             data=KM_Righting)
region_righting_dif

#Cox proportional hazards model fit
region_righting_cox=coxph(Surv(Duration,Status)~Block+Region+Sex+CW, 
                          data=KM_Righting)
summary(region_righting_cox)

#ANOVA of main effects 
anova(region_righting_cox)
```


Analysis of Presentation Figures
--
The following is more in depth analysis for each publication figure generated to illustrate the main effects (see Righting Response Figures.rmd)


Figure 1: Sublethal Site
Kaplan Meier curve of righting probability in P. armatus from both study sites at each location; from North to South, Baruch, Charleston, Bluffton (SC) and Savannah (GA) during cold temperature trial. Log-Rank test: X2 = 20.4, p = 0.617
```{r, echo=F}
#Log Rank test 
survdiff(Surv(Duration,Status)~Block+Site, data=KM_Righting)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Site, data=KM_Righting))
#ANCOVA of main effects 
anova(coxph(Surv(Duration,Status)~Block+Site, data=KM_Righting))

```

Figure 2: Sublethal Region
Kaplan Meier curve of righting probability in P. armatus from Northern and Southern sites during cold temperature trial. Log-Rank test: X2 = 10.4, p = 0.064
```{r, echo=F}
#Log Rank test 
survdiff(Surv(Duration,Status)~Block+Region, data=KM_Righting)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Region, data=KM_Righting))
```

Figure 3: Sublethal Sex
Kaplan Meier curve of righting probability in P. armatus females and males during cold temperature trial. Log-Rank test: X2 = 15.9, p < 0.01
```{r, echo=F}
#Log Rank test 
survdiff(Surv(Duration,Status)~Block+Sex, data=KM_Righting)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Sex, data=KM_Righting))
```

Figure 4: Sublethal Sex by Region
Kaplan Meier curve of righting probability in P. armatus females and males from Northern or Southern sites during cold temperature trial. Log-Rank test: X2 = 22.1, p < 0.05
```{r, echo=F}
#Log Rank
survdiff(Surv(Duration,Status)~Block+Sex+Region, data=KM_Righting)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Sex+Region, data=KM_Righting))
```

Figure 5: Sublethal Size Regression
Carapace width (mm) of P. armatus plotted by time of righting response loss for females (R2 = 0.036, p <0.05) and males (R2 = 0.062, p < 0.01 ) during cold temperature trial. 
```{r, echo=F}
#subset
lethal_males=subset(KM_Righting, Sex == 'Male')
lethal_females=subset(KM_Righting, Sex == 'Female')

#Regression
summary(lm(CW~Duration, data=lethal_males))
summary(lm(CW~Duration, data=lethal_females))
