---
title: "Applied Quant Analysis"
author: "Kevin Mack"
date: "3/25/2016"
output: pdf_document
---

Preliminary Analysis
--
This is preliminary analysis of lethal and righting responses of P. armatus. I ran an ANCOVA and on the main effects: site, sex, and block with crab sixe (CW) as a covariate. A significant block effect was detected. This prelimary analysis indicates that Sex was the strongest effect, followed by the covariate Carapace Width, and Site. There were no significant interactions.
```{r}
KM_Survival <- read.csv("~/Incubation Analysis/Trial1_survival.csv")
KM_Righting <- read.csv("~/Incubation Analysis/Trial1_sublethal.csv")
KM_Survival$Block=as.factor(KM_Survival$Block)
KM_Righting$Block=as.factor(KM_Righting$Block)
library(survival)

#ANCOVA of main effects: Lethal
options(contrasts=c("contr.sum", "contr.poly"))
fit1=(aov(lm(Duration~Site*Sex*Block+CW, data=KM_Survival)))
drop1(fit1,~.,test='F')

#Boxplots: Lethal
plot(KM_Survival$Sex, KM_Survival$Duration)
plot(KM_Survival$Site, KM_Survival$Duration)
plot(KM_Survival$CW,KM_Survival$Duration)
plot(KM_Survival$Duration, KM_Survival$CW ) #closley related to sex. Examine males and females separetely?

#ANCOVA of main effests: Righting Response
options(contrasts=c("contr.sum", "contr.poly"))
fit1=(aov(lm(Duration~Site*Sex*Block+CW, data=KM_Righting)))
drop1(fit1,~.,test='F')

#Boxplots: Righting Response
plot(KM_Righting$Sex, KM_Righting$Duration)
plot(KM_Righting$Site, KM_Righting$Duration)
plot(KM_Righting$CW,KM_Righting$Duration)
plot(KM_Righting$Duration, KM_Righting$CW ) #closley related to sex. Examine males and females separetely?


```

Analysis of Main Effects
--
This is analysis of lethal and righting responses of P. armatus. I tested the main effects: site, sex, block and crab sixe (CW) by generating Kaplan-Meier curves for lethal and righting response data, main effects were tested with a Log Rank test. I next fitted these curves to a Cox proportional hazards model to determine differences in hazard risk between main effects, and which were tested with ANOVA.

```{r}
#LETHAL-------------------------------------------------------------------------
#KM Curves of main effects
let_mod1=survfit(Surv(Duration,Status)~Block+Site+Sex, data=KM_Survival)
plot(let_mod1, ylab='Survival Probability', xlab='Duration') #disregard everything this figure is not for publication

#Log Rank test of main effects tests for differences
let_mod1_dif=survdiff(Surv(Duration,Status)~Block+Site+Sex, data=KM_Survival)
let_mod1_dif

#Cox proportional hazards model fit
let_mod1_cox=coxph(Surv(Duration,Status)~Block+Site+Sex+CW, data=KM_Survival)
summary(let_mod1_cox)

#ANOVA of main effects 
anova(let_mod1_cox)

#RIGHTING RESPONSE------------------------------------------------------------
#KM Curves of main effects
rr_mod1=survfit(Surv(Duration,Status)~Block+Site+Sex, data=KM_Righting)
plot(rr_mod1, ylab='Survival Probability', xlab='Duration') #disregard everything this figure is not for publication

#Log Rank test of main effects tests for differences
rr_mod1_dif=survdiff(Surv(Duration,Status)~Block+Site+Sex, data=KM_Righting)
rr_mod1_dif

#Cox proportional hazards model fit
rr_mod1_cox=coxph(Surv(Duration,Status)~Block+Site+Sex+CW, data=KM_Righting)
summary(rr_mod1_cox)

#ANOVA of main effects 
anova(rr_mod1_cox)
```




Spatial Analysis
--
Analysis of lethal and righting responses of P. armatus at varying spatial scales: Site, Location, Region

Analysis by SITE
```{r}
#LETHAL
#KM Curves of main effects
site=survfit(Surv(Duration,Status)~Block+Site+Sex, data=KM_Survival)
plot(site, ylab='survival probabilty', xlab='Duration', main='Model 1') #disregard everything

#Log Rank test of main effects tests for differences
site_dif=survdiff(Surv(Duration,Status)~Block+Site+Sex, data=KM_Survival)
site_dif

#Cox proportional hazards model fit
site_cox=coxph(Surv(Duration,Status)~Block+Site+Sex+CW, data=KM_Survival)
summary(site_cox)

#ANCOVA of main effects 
anova(site_cox)

#SUBLETHAL
#KM Curves of main effects
site_righting=survfit(Surv(Duration,Status)~Block+Site+Sex, data=KM_Righting)
plot(site_righting, ylab='survival probabilty', xlab='Duration', main='Model 1') #disregard everything

#Log Rank test of main effects tests for differences
site_righting_dif=survdiff(Surv(Duration,Status)~Block+Site+Sex, data=KM_Righting)
site_righting_dif

#Cox proportional hazards model fit
site_righting_cox=coxph(Surv(Duration,Status)~Block+Site+Sex+CW, data=KM_Righting)
summary(site_righting_cox)

#ANCOVA of main effects 
anova(site_righting_cox)

```


Analysis by LOCATION
```{r}
#LETHAL
#KM Curves of main effects
location=survfit(Surv(Duration,Status)~Block+Location+Sex, data=KM_Survival)
plot(location, ylab='righting probabilty', xlab='Duration', main='Model 1') #disregard everything

#Log Rank test of main effects tests for differences
location_dif=survdiff(Surv(Duration,Status)~Block+Location+Sex, data=KM_Survival)
location_dif

#Cox proportional hazards model fit
location_cox=coxph(Surv(Duration,Status)~Block+Location+Sex+CW, data=KM_Survival)
summary(location_cox)

#ANCOVA of main effects 
anova(location_cox)

#SUBLETHAL
#KM Curves of main effects
location_righting=survfit(Surv(Duration,Status)~Block+Location+Sex, data=KM_Righting)
plot(location_righting, ylab='righting probabilty', xlab='Duration', main='Model 1') #disregard everything

#Log Rank test of main effects tests for differences
location_righting_dif=survdiff(Surv(Duration,Status)~Block+Location+Sex, data=KM_Righting)
location_righting_dif

#Cox proportional hazards model fit
location_righting_cox=coxph(Surv(Duration,Status)~Block+Location+Sex+CW, data=KM_Righting)
summary(location_righting_cox)

#ANCOVA of main effects 
anova(location_righting_cox)
```


Analysis by REGION
```{r}
#LETHAL
#KM Curves of main effects
region=survfit(Surv(Duration,Status)~Block+Region+Sex, data=KM_Survival)
plot(region, ylab='survival probabilty', xlab='Duration', main='Model 1')
#disregard everything

#Log Rank test of main effects tests for differences
region_dif=survdiff(Surv(Duration,Status)~Block+Region+Sex, data=KM_Survival)
region_dif

#Cox proportional hazards model fit
region_cox=coxph(Surv(Duration,Status)~Block+Region+Sex+CW, data=KM_Survival)
summary(region_cox)

#ANOVA of main effects 
anova(region_cox)

#SUBLETHAL
#KM Curves of main effects
region_righting=survfit(Surv(Duration,Status)~Block+Region+Sex, data=KM_Righting)
plot(region_righting, ylab='survival probabilty', xlab='Duration', main='Model 1')
#disregard everything

#Log Rank test of main effects tests for differences
region_righting_dif=survdiff(Surv(Duration,Status)~Block+Region+Sex, data=KM_Righting)
region_righting_dif

#Cox proportional hazards model fit
region_righting_cox=coxph(Surv(Duration,Status)~Block+Region+Sex+CW, data=KM_Righting)
summary(region_righting_cox)

#ANOVA of main effects 
anova(region_righting_cox)
```

How to decide what to present? 

