---
title: "BEM_RightingResponse"
author: "Kevin Mack"
date: "2/18/2016"
output: pdf_document
---

Preliminary Analysis
--
This is preliminary/esploratory analysis of sublethal righting responses of P. armatus. I ran an ANCOVA and on the main effects: site, sex, and block with crab size (CW) as a covariate. A significant block effect was detected. This prelimary analysis indicates that the covariate CW was the strongest effect. There were no significant interactions.
```{r}
KM_Righting <- read.csv("~/Applied Quant Project/Trial1_sublethal.csv")
KM_Righting$Block=as.factor(KM_Righting$Block)
library(survival)

#ANCOVA of main effests: Righting Response
options(contrasts=c("contr.sum", "contr.poly"))
fit1=(aov(lm(Duration~Site*Sex*Block+CW, data=KM_Righting)))
drop1(fit1,~.,test='F')

#Boxplots: Righting Response
plot(KM_Righting$Sex, KM_Righting$Duration)
plot(KM_Righting$Site, KM_Righting$Duration)
plot(KM_Righting$CW,KM_Righting$Duration)
plot(KM_Righting$Duration, KM_Righting$CW ) 
#closley related to sex. Examine males and females separetely?

```


Analysis of Main Effects
--
This is analysis of righting responses of P. armatus. I tested the main effects: site, sex, block and crab size (CW) by generating Kaplan-Meier curves for righting response data, main effects were tested with a Log Rank test. I next fitted these curves to a Cox proportional hazards model to determine differences in hazard risk between main effects, which were tested with ANOVA.
```{r}
#KM Curves of main effects
rr_mod1=survfit(Surv(Duration,Status)~Block+Site+Sex, data=KM_Righting)
plot(rr_mod1, ylab='Survival Probability', xlab='Duration') 
#disregard this figure is not for publication

#Log Rank test of main effects tests for differences
rr_mod1_dif=survdiff(Surv(Duration,Status)~Block+Site+Sex, data=KM_Righting)
rr_mod1_dif

#Cox proportional hazards model fit
rr_mod1_cox=coxph(Surv(Duration,Status)~Block+Site+Sex+CW, data=KM_Righting)
summary(rr_mod1_cox)

#ANOVA of main effects 
anova(rr_mod1_cox)
```

Analysis of Spatial Scale
--
The following is analysis of righting responses of P. armatus at three spatial scales. I tested the main effects, grouping data by site, location, and region, to determine inter-site and region differences in righting response. Differences in righting response exist between sites, locations, and regions.
```{r}
##Analysis by SITE
#KM Curves of main effects
site_righting=survfit(Surv(Duration,Status)~Block+Site+Sex, data=KM_Righting)
plot(site_righting, ylab='survival probabilty', xlab='Duration', main='Model 1') 
#disregard

#Log Rank test of main effects tests for differences
site_righting_dif=survdiff(Surv(Duration,Status)~Block+Site+Sex,
                           data=KM_Righting)
site_righting_dif

#Cox proportional hazards model fit
site_righting_cox=coxph(Surv(Duration,Status)~Block+Site+Sex+CW, 
                        data=KM_Righting)
summary(site_righting_cox)

#ANCOVA of main effects 
anova(site_righting_cox)
```

```{r}
##Analysis by LOCATION
#KM Curves of main effects
location_righting=survfit(Surv(Duration,Status)~Block+Location+Sex, 
                          data=KM_Righting)
plot(location_righting, ylab='righting probabilty', xlab='Duration', 
     main='Model 1') #disregard 

#Log Rank test of main effects tests for differences
location_righting_dif=survdiff(Surv(Duration,Status)~Block+Location+Sex,
                               data=KM_Righting)
location_righting_dif

#Cox proportional hazards model fit
location_righting_cox=coxph(Surv(Duration,Status)~Block+Location+Sex+CW,
                            data=KM_Righting)
summary(location_righting_cox)

#ANCOVA of main effects 
anova(location_righting_cox)
```

```{r}
##Analysis by REGION
#KM Curves of main effects
region_righting=survfit(Surv(Duration,Status)~Block+Region+Sex, 
                        data=KM_Righting)
plot(region_righting, ylab='survival probabilty', xlab='Duration', 
     main='Model 1')
#disregard

#Log Rank test of main effects tests for differences
region_righting_dif=survdiff(Surv(Duration,Status)~Block+Region+Sex, 
                             data=KM_Righting)
region_righting_dif

#Cox proportional hazards model fit
region_righting_cox=coxph(Surv(Duration,Status)~Block+Region+Sex+CW, 
                          data=KM_Righting)
summary(region_righting_cox)

#ANOVA of main effects 
anova(region_righting_cox)
```


Analysis of Presentation Figures
--
The following is analysis of each of the presentation figures
Figure 6: Sublethal Site
```{r, echo=F}
#Log Rank test 
survdiff(Surv(Duration,Status)~Block+Site, data=KM_Righting)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Site, data=KM_Righting))
#ANCOVA of main effects 
anova(coxph(Surv(Duration,Status)~Block+Site, data=KM_Righting))

```

Figure 7: Sublethal Region
```{r, echo=F}
#Log Rank test 
survdiff(Surv(Duration,Status)~Block+Region, data=KM_Righting)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Region, data=KM_Righting))
#ANCOVA of main effects 
anova(coxph(Surv(Duration,Status)~Block+Region, data=KM_Righting))
```

Figure 8: Sublethal Sex
```{r, echo=F}
#Log Rank test 
survdiff(Surv(Duration,Status)~Block+Sex, data=KM_Righting)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Sex, data=KM_Righting))
#ANCOVA of main effects 
anova(coxph(Surv(Duration,Status)~Block+Sex, data=KM_Righting))
```

Figure 9: Sublethal Sex by Region
```{r, echo=F}
#Log Rank
survdiff(Surv(Duration,Status)~Block+Sex+Region, data=KM_Righting)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Sex+Region, data=KM_Righting))
#ANOVA
anova(coxph(Surv(Duration,Status)~Block+Sex+Region, data=KM_Righting))
```

Figure 10: Sublethal Size Regression
```{r, echo=F}
#subset
lethal_males=subset(KM_Righting, Sex == 'Male')
lethal_females=subset(KM_Righting, Sex == 'Female')

#Regression
summary(lm(CW~Duration, data=lethal_males))
summary(lm(CW~Duration, data=lethal_females))
