---
title: "Survival Analysis"
author: "Kevin Mack"
date: "3/30/2016"
output: pdf_document
---

Preliminary Analysis
--
The following is preliminary/exploratory analysis of lethal responses of P. armatus. I ran an ANCOVA and on the main effects: site, sex, and block with crab size (CW) as a covariate. A significant block effect was detected. This prelimary analysis indicates that Sex was the strongest effect, followed by the covariate Carapace Width, and Site. There were no significant interactions.
```{r}
KM_Survival <- read.csv("~/Applied Quant Project/Trial1_survival.csv")
KM_Survival$Block=as.factor(KM_Survival$Block)
library(survival)

#ANCOVA of main effects: Lethal
options(contrasts=c("contr.sum", "contr.poly"))
fit1=(aov(lm(Duration~Site*Sex*Block+CW, data=KM_Survival)))
drop1(fit1,~.,test='F')

#Boxplots: Lethal
plot(KM_Survival$Sex, KM_Survival$Duration)
plot(KM_Survival$Site, KM_Survival$Duration)
plot(KM_Survival$CW,KM_Survival$Duration)
plot(KM_Survival$Duration, KM_Survival$CW ) 
#closley related to sex. Examine males and females separetely?
```

Analysis of Main Effects
--
The following is analysis of lethal responses of P. armatus. I tested the main effects: site, sex, block and crab size (CW) by generating Kaplan-Meier curves, main effects were tested with a Log Rank test. I next fitted these curves to a Cox proportional hazards model to determine differences in hazard risk between main effects, which were tested with ANOVA. A significant block effect was detected. Significant differences in survival were detected for all the main effects.
```{r}
#KM Curves of main effects
let_mod1=survfit(Surv(Duration,Status)~Block+Site+Sex, data=KM_Survival)
plot(let_mod1, ylab='Survival Probability', xlab='Duration') 
#disregard, this figure is not for publication

#Log Rank test of main effects tests for differences
let_mod1_dif=survdiff(Surv(Duration,Status)~Block+Site+Sex, data=KM_Survival)
let_mod1_dif

#Cox proportional hazards model fit
let_mod1_cox=coxph(Surv(Duration,Status)~Block+Site+Sex+CW, data=KM_Survival)
summary(let_mod1_cox)

#ANOVA of main effects 
anova(let_mod1_cox)
```

Analysis of Spatial Scale
--
The following is analysis of lethal responses of P. armatus at three spatial scales. I tested the main effects, grouping data by site, location, and region, to determine inter-site and region differences in survival. Differences in survival exist between sites, and regions, but were not detcted between locations.
```{r}
##Analysis by SITE
#KM Curves of main effects
let_site=survfit(Surv(Duration,Status)~Block+Site+Sex, data=KM_Survival)
plot(let_site, ylab='survival probabilty', xlab='Duration', main='Model 1') 
#disregard figure

#Log Rank test of main effects tests for differences
let_site_dif=survdiff(Surv(Duration,Status)~Block+Site+Sex, data=KM_Survival)
let_site_dif

#Cox proportional hazards model fit
let_site_cox=coxph(Surv(Duration,Status)~Block+Site+Sex+CW, data=KM_Survival)
summary(let_site_cox)

#ANCOVA of main effects 
anova(let_site_cox)
```

```{r}
##Analysis by LOCATION
#KM Curves of main effects
let_location=survfit(Surv(Duration,Status)~Block+Location+Sex, data=KM_Survival)
plot(let_location, ylab='righting probabilty', xlab='Duration', main='Model 1')
#disregard everything

#Log Rank test of main effects tests for differences
let_location_dif=survdiff(Surv(Duration,Status)~Block+Location+Sex, 
                      data=KM_Survival)
let_location_dif

#Cox proportional hazards model fit
let_location_cox=coxph(Surv(Duration,Status)~Block+Location+Sex+CW, 
                   data=KM_Survival)
summary(let_location_cox)

#ANCOVA of main effects 
anova(let_location_cox)
```

```{r}
##Analysis by REGION
#KM Curves of main effects
let_region=survfit(Surv(Duration,Status)~Block+Region+Sex, data=KM_Survival)
plot(let_region, ylab='survival probabilty', xlab='Duration', main='Model 1')
#disregard everything

#Log Rank test of main effects tests for differences
let_region_dif=survdiff(Surv(Duration,Status)~Block+Region+Sex, data=KM_Survival)
let_region_dif

#Cox proportional hazards model fit
let_region_cox=coxph(Surv(Duration,Status)~Block+Region+Sex+CW, data=KM_Survival)
summary(let_region_cox)

#ANOVA of main effects 
anova(let_region_cox)
```


Analysis of Presentation Figures
--
The following is analysis of each of the presentation figures

Figure 1: Lethal Site
```{r, echo=F}
#Log Rank
survdiff(Surv(Duration,Status)~Block+Site, data=KM_Survival)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Site, data=KM_Survival))
#ANCOVA
anova(coxph(Surv(Duration,Status)~Block+Site, data=KM_Survival))
```

Figure 2: Lethal Region
```{r, echo=F}
#Log Rank
survdiff(Surv(Duration,Status)~Block+Region, data=KM_Survival)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Region, data=KM_Survival))
#ANOVA
anova(coxph(Surv(Duration,Status)~Block+Region, data=KM_Survival))
```

Figure 3: Lethal Sex
```{r, echo=F}
#Log Rank
survdiff(Surv(Duration,Status)~Block+Sex, data=KM_Survival)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Sex, data=KM_Survival))
#ANOVA
anova(coxph(Surv(Duration,Status)~Block+Sex, data=KM_Survival))
```

Figure 4: Lethal Sex by Region
```{r, echo=F}
#Log Rank
survdiff(Surv(Duration,Status)~Block+Sex+Region, data=KM_Survival)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Sex+Region, data=KM_Survival))
#ANOVA
anova(coxph(Surv(Duration,Status)~Block+Sex+Region, data=KM_Survival))
```

Figure 5: Lethal Size Regression
```{r, echo=F}
#subset
lethal_males=subset(KM_Survival, Sex == 'Male')
lethal_females=subset(KM_Survival, Sex == 'Female')

#Regression
summary(lm(CW~Duration, data=lethal_males))
summary(lm(CW~Duration, data=lethal_females))
```





