---
title: "Survival Analysis"
author: "Kevin Mack"
date: "3/30/2016"
output: pdf_document
---

Preliminary Analysis
--
The following is preliminary/exploratory analysis of lethal responses of _P. armatus_. I ran an ANCOVA and on the main effects: site, sex, and block with crab size (CW) as a covariate. A significant block effect was detected. This prelimary analysis indicates that Sex was the strongest effect, followed by the covariate CW, and Site. There were no significant interactions. This analysis informed the next analysis of main effects.
```{r}
KM_Survival <- read.csv("~/Applied Quant Project/Survival/Trial1_survival.csv")
KM_Survival$Block=as.factor(KM_Survival$Block)
library(survival)

#ANCOVA of main effects: Lethal
options(contrasts=c("contr.sum", "contr.poly"))
fit1=(aov(lm(Duration~Site*Sex*Block+CW, data=KM_Survival)))
drop1(fit1,~.,test='F')

#Boxplots: Lethal
plot(KM_Survival$Sex, KM_Survival$Duration)
plot(KM_Survival$Site, KM_Survival$Duration)
plot(KM_Survival$CW,KM_Survival$Duration)
plot(KM_Survival$Duration, KM_Survival$CW ) 
#closley related to sex. Examine males and females separetely?
```

Analysis of Main Effects
--
The following is analysis of lethal responses of _P. armatus_. I tested the main effects: site, sex, block and crab size (CW) by generating Kaplan-Meier curves, main effects were tested with a Log Rank test. I next fitted these curves to a Cox proportional hazards model to determine differences in hazard risk between main effects, which were further summarized with ANOVA. A significant block effect was detected. Significant differences in survival were detected for all the main effects.
```{r}
#KM Curves of main effects
let_mod1=survfit(Surv(Duration,Status)~Block+Site+Sex, data=KM_Survival)
plot(let_mod1, ylab='Survival Probability', xlab='Duration') 

#Log Rank test of main effects tests for differences
let_mod1_dif=survdiff(Surv(Duration,Status)~Block+Site+Sex, data=KM_Survival)
let_mod1_dif

#Cox proportional hazards model fit
let_mod1_cox=coxph(Surv(Duration,Status)~Block+Site+Sex+CW, data=KM_Survival)
summary(let_mod1_cox)

#ANOVA of main effects 
anova(let_mod1_cox)
```

Analysis of Spatial Scale
--
The following is analysis of lethal responses of _P. armatus_ at three spatial scales. I tested the main effects, grouping data by site, location, and region, to determine inter-site and region differences in survival. Differences in survival exist between sites, and regions, but were not detcted between locations. Based on this analysis data can be summarized by region, North vs South.
```{r}
##Analysis by SITE
#KM Curves of main effects
let_site=survfit(Surv(Duration,Status)~Block+Site+Sex, data=KM_Survival)
plot(let_site, ylab='survival probabilty', xlab='Duration', main='Site') 

#Log Rank test of main effects tests for differences
let_site_dif=survdiff(Surv(Duration,Status)~Block+Site+Sex, data=KM_Survival)
let_site_dif

#Cox proportional hazards model fit
let_site_cox=coxph(Surv(Duration,Status)~Block+Site+Sex+CW, data=KM_Survival)
summary(let_site_cox)

#ANCOVA of main effects 
anova(let_site_cox)
```

```{r}
##Analysis by LOCATION
#KM Curves of main effects
let_location=survfit(Surv(Duration,Status)~Block+Location+Sex, data=KM_Survival)
plot(let_location, ylab='righting probabilty', xlab='Duration', main='Location')

#Log Rank test of main effects tests for differences
let_location_dif=survdiff(Surv(Duration,Status)~Block+Location+Sex, 
                      data=KM_Survival)
let_location_dif

#Cox proportional hazards model fit
let_location_cox=coxph(Surv(Duration,Status)~Block+Location+Sex+CW, 
                   data=KM_Survival)
summary(let_location_cox)

#ANCOVA of main effects 
anova(let_location_cox)
```

```{r}
##Analysis by REGION
#KM Curves of main effects
let_region=survfit(Surv(Duration,Status)~Block+Region+Sex, data=KM_Survival)
plot(let_region, ylab='survival probabilty', xlab='Duration', main='Region')

#Log Rank test of main effects tests for differences
let_region_dif=survdiff(Surv(Duration,Status)~Block+Region+Sex, data=KM_Survival)
let_region_dif

#Cox proportional hazards model fit
let_region_cox=coxph(Surv(Duration,Status)~Block+Region+Sex+CW, data=KM_Survival)
summary(let_region_cox)

#ANOVA of main effects 
anova(let_region_cox)
```


Analysis of Presentation Figures
--
The following is analysis of each of the presentation figures

Figure 1: Lethal Site
Kaplan Meier curve of survival probability in _P. armatus_ from both study sites at each location; from North to South, Baruch, Charleston, Bluffton (SC) and Savannah (GA) during cold temperature trial. Log-Rank test: X2 = 55.6, p < 0.001.
```{r, echo=F}
#Log Rank
survdiff(Surv(Duration,Status)~Block+Site, data=KM_Survival)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Site, data=KM_Survival))
```

Figure 2: Lethal Region
Kaplan Meier curve of survival probability in _P. armatus_ from Northern and Southern sites during cold temperature trial. Log-Rank test: X2 = 31.5, p < 0.001.
```{r, echo=F}
#Log Rank
survdiff(Surv(Duration,Status)~Block+Region, data=KM_Survival)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Region, data=KM_Survival))
```

Figure 3: Lethal Sex
Kaplan Meier curve of survival probability in _P. armatus_ females and males during cold temperature trial. Log-Rank test: X2 = 54.9, p < 0.001.
```{r, echo=F}
#Log Rank
survdiff(Surv(Duration,Status)~Block+Sex, data=KM_Survival)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Sex, data=KM_Survival))
```

Figure 4: Lethal Sex by Region
Kaplan Meier curve of survival probability in P. armatus females and males from Northern and Southern sites during cold temperature trial. Log-Rank test: X2 = 61.6, p < 0.001.
```{r, echo=F}
#Log Rank
survdiff(Surv(Duration,Status)~Block+Sex+Region, data=KM_Survival)
#Cox proportional hazards model fit
summary(coxph(Surv(Duration,Status)~Block+Sex+Region, data=KM_Survival))
```

Figure 5: Lethal Size Regression
Carapace width (mm) of _P. armatus_ plotted by time of righting response loss for females (R2 = 0.00, p = 0.88) and males (R2 = 0.10, p < 0.001) during cold temperature trial.
```{r, echo=F}
#subset
lethal_males=subset(KM_Survival, Sex == 'Male')
lethal_females=subset(KM_Survival, Sex == 'Female')

#Regression
summary(lm(CW~Duration, data=lethal_males))
summary(lm(CW~Duration, data=lethal_females))
```





